Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-06 Grassmannian_cells.tex
Some section 3 edits

Write $\Sigma_1$ and $\Sigma_2$ for the BGP-reflection functors of $K(n)$ \cite{bgp}. 
Note that (up to relabeling vertices) the quiver $K(n)$ itself does not change when applying the reflection functors and thus we can view these as endofunctors acting on the category $\rep K(n)$.

The endofunctors $\Sigma_1^2$ and $\Sigma_2^2$ are each naturally isomorphic to the identity maps on the full subcategories $\rep_{\langle 1\rangle} K(n),\rep_{\langle 2\rangle} K(n)\subset \rep K(n)$, where the objects of $\rep_{\langle i\rangle} K(n)$ are those representations of $K(n)$ which do not contain the simple $S_i$ as a direct summand.
Also, following \cite{brenner-butler}, the Auslander-Reiten translation $\tau:\rep K(n)\to\rep K(n)$ may be identified with the functor $\Sigma_2\Sigma_1$.
\sayD{These edits make the statements false.  By viewing the reflection functors as endofunctors, we have to interchange the vertex labels so that the inverse of $\Sigma_1$ is $\Sigma_2$.  Maybe we want to explicitly introduce the functor identifying $\rep K(n)$ with $\rep K(n)^{op}$, is there a standard notation for this?}

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
The following is well-known.
\begin{theorem}
  \label{th:rigids}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  \begin{enumerate}
    \item These representations satisfy the following recursions using the reflection functors:
      \[P_1=S_1,\quad P_m=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
      for $m\ge2$.
      In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$.
    \item If $\ses{M}{B}{N}$ is a short exact sequence such that no direct summand of $M$, $B$ and $N$ is preinjective, then $\ses{(\Sigma_1\Sigma_2)^nM}{(\Sigma_1\Sigma_2)^nB}{(\Sigma_1\Sigma_2)^nN}$ and $\ses{\Sigma_2(\Sigma_1\Sigma_2)^nM}{\Sigma_2(\Sigma_1\Sigma_2)^nB}{\Sigma_2(\Sigma_1\Sigma_2)^nN}$ are short exact sequences without any preinjective direct summands.
      Without the assumption on the direct summands, it is not clear that short exact sequences are preserved.
      This is because $\Sigma_2 S_2=0$.
  \end{enumerate}
\end{remark}
We will see that the indecomposable representations of dimension $\alpha(m,r):=\dim P_{m+1}-r\dim P_m$ with $0\leq r\leq n-1$ play an important role as they appear as quotients of morphisms between preprojective representations.
\begin{proposition}
  \label{pro:indecomposables}
  Let $m\geq 1$. Let $R_\alpha(K(n))$ be the affine space of representations of dimension $\alpha(m,r)$.
  The following hold:
  %\sayD{Move the first two of these into section~\ref{sec:truncated preprojectives}?}
  \begin{enumerate}
    \item The isomorphism classes of indecomposable representations of dimension $\alpha(m,r)$ are in one-to-one correspondence to points of $\Gr_{n-r}(\CC^n)$.
    \item The set indecomposable representations of dimension $\alpha(m,r)$ is given by a non-empty open subset of $R_\alpha(K(n))$. 
   % \item The action of $\Gl_n(\CC)$ is transitive on $\mathrm{Ind}(K(n),\alpha(m,r))$.
  \end{enumerate}
\end{proposition}
\begin{proof}

As the reflection functors $\Sigma_1,\,\Sigma_2$ preserve indecomposability and as we have $\Sigma_2(\alpha(m,r))=\alpha(m+1,r)$, it suffices to prove the first statement for $m=1$. Then we have $\alpha(m,r)=(1,d)$ for $d:=n-r$. This means that $M\in R_\alpha(K(n))$ is given by a matrix $M\in\CC^{d\times n}$ where the $i^{\mathrm{th}}$ column stands for $M_{a_i}$. Now $M$ is indecomposable if and only if $\rk{M}=d$. Indeed, it is clear that $S_1^{m}$ splits up if and only if $\rk{M}=d-m$.

This shows that the indecomposable representations in $R_{\alpha(1,d)}(K(n))$  are in one-to-one correspondence with matrices of maximal rank. Thus they are in one-to-one correspondence to subspaces of $\CC^n$ of dimension $d$ spanned by the column vectors of $M$. Now it straightforward to check that the $\Gl_\alpha=\CC^\ast\times\Gl_d(\CC)$-action on $R_\alpha(K(n))$ corresponds to the base change action of $\Gl_d(\CC)$ on the set of subspaces. This shows the first statement.

@@

%For $m=1$, the action of $\Gl_n(\CC)$ on $\Gr_d(\CC^n)$ is transitive which shows that it is transitive on $\Ind_\alpha(K(n))$. As the reflection functors preserve indecomposability and isomorphism classes, by Lemma \ref{compatibleGSigma}, for every $g\in\Gl_n(\CC)$, we obtain a commutative diagram
%\[\xymatrix{\Ind(K(n),\alpha(m,r))\ar[r]^{\Sigma_2}\ar[d]^{g^{-T}}&\Ind(K(n),\alpha(m+1,r))\ar[d]^g\\
%\Ind(K(n),\alpha(m,r))\ar[r]^{\Sigma_2}&\Ind(K(n),\alpha(m+1,r))}\]
%As the $\Gl_n(\CC)$-action is transitive on the left hand side, this implies that it is also transitive on the right hand side.
\end{proof}
\begin{remark}There is a more elegant way to prove Proposition \ref{pro:indecomposables} using the notion of stability and moduli spaces. Actually, fixing the standard stability induced by the linear form $\Theta(\alpha)=\alpha_2$, it can be shown that all indecomposables are stable and that the moduli space of stable representation is in fact $\Gr_d(\CC^n)$. We inserted this proof because the notion of stability were only used at this point and we wanted to keep this part as simple as possible. 
\end{remark}

>>>>

Write $\Sigma_1$ and $\Sigma_2$ for the BGP-reflection functors of $K(n)$ \cite{bgp}. 
We use the same symbols $\Sigma_1$, $\Sigma_2$ for the BGP-reflection functors of $K(n)^{op}$, this should not lead to any confusion. 
Then the endofunctors $\Sigma_1^2$ and $\Sigma_2^2$ are each naturally isomorphic to the identity maps on the full subcategories $\rep_{\langle 1\rangle} K(n),\rep_{\langle 2\rangle} K(n)\subset \rep K(n)$, where the objects of $\rep_{\langle i\rangle} K(n)$ are those representations of $K(n)$ which do not contain the simple $S_i$ as a direct summand.
Also, following \cite{brenner-butler}, the Auslander-Reiten translation $\tau:\rep K(n)\to\rep K(n)$ may be identified with the functor $\Sigma_2\Sigma_1$.

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
The following is well-known.
\begin{theorem}
  \label{th:rigids}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  We may identify the quiver $K(n)$ with $K(n)^{op}$ by interchanging the vertex labels.
  This induces an isomorphism of categories $\rep K(n)\cong\rep K(n)^{op}$ which we write as $M\mapsto M^\sigma$.
  Note that $\Sigma_1(M^\sigma)=(\Sigma_2 M)^\sigma$ and $\Sigma_2(M^\sigma)=(\Sigma_1 M)^\sigma$.
  \begin{enumerate}
    \item The preprojective and preinjective representations satisfy the following recursions using the reflection functors:
      \[P_1=S_1,\quad P_m^\sigma=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m^\sigma=\Sigma_1 I_{m-1}\]
      for $m\ge2$.
      In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$.
    \item If $\ses{M}{B}{N}$ is a short exact sequence such that no direct summand of $M$, $B$, nor $N$ is preinjective, then $\ses{(\Sigma_1\Sigma_2)^nM}{(\Sigma_1\Sigma_2)^nB}{(\Sigma_1\Sigma_2)^nN}$ and $\ses{\Sigma_2(\Sigma_1\Sigma_2)^nM}{\Sigma_2(\Sigma_1\Sigma_2)^nB}{\Sigma_2(\Sigma_1\Sigma_2)^nN}$ are short exact sequences without any preinjective direct summands.
      %Without the assumption on the direct summands, it is not clear that short exact sequences are preserved.
      %This is because $\Sigma_2 S_2=0$.
  \end{enumerate}
\end{remark}
We will see that the indecomposable representations of dimension $\alpha(m,r):=\dim P_{m+1}-r\dim P_m$ with $0\leq r\leq n-1$ play an important role as they appear as quotients of morphisms between preprojective representations.
For a dimension vector $\alpha$, write $R_\alpha(K(n))$ for the affine space of representations of dimension $\alpha$.
\begin{proposition}
  \label{pro:indecomposables}
  Let $m\geq 1$ and $0\le r\le n-1$.
  The following hold:
  %\sayD{Move the first two of these into section~\ref{sec:truncated preprojectives}?}
  \begin{enumerate}
    \item The isomorphism classes of indecomposable representations of dimension $\alpha(m,r)$ are in one-to-one correspondence to points of $\Gr_{n-r}(\CC^n)$.
    \item The set of indecomposable representations of dimension $\alpha(m,r)$ is given by a non-empty open subset of $R_{\alpha(m,r)}(K(n))$. 
   % \item The action of $\Gl_n(\CC)$ is transitive on $\mathrm{Ind}(K(n),\alpha(m,r))$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  As the reflection functors $\Sigma_1,\,\Sigma_2$ preserve indecomposability and as we have $\Sigma_2(\alpha(m,r))=\alpha(m+1,r)$, it suffices to prove the first statement for $m=1$.
  Then we have $\alpha(m,r)=(1,d)$ for $d:=n-r$.
  This means that $X\in R_\alpha(K(n))$ is given by a matrix $M_X\in\CC^{d\times n}$, where the $i^{\mathrm{th}}$ column stands for $X_{\alpha_i}$.
  Now $X$ is indecomposable if and only if $\rk(M_X)=d$.
  Indeed, $X$ admits a summand isomorphic to $S_1^m$ exactly when $\rk(M_X)=d-m$.

  This shows that the indecomposable representations in $R_{\alpha(1,d)}(K(n))$ are in one-to-one correspondence with $d\times n$ matrices of maximal rank.
  Thus they are in one-to-one correspondence to subspaces of $\CC^n$ of dimension $d$ spanned by the column vectors of $M_X$.
  Now it is straightforward to check that the $\Gl_\alpha=\CC^\ast\times\Gl_d(\CC)$-action on $R_\alpha(K(n))$ corresponds to the base change action of $\Gl_d(\CC)$ on the set of subspaces.
  This shows the first statement.

@@

%For $m=1$, the action of $\Gl_n(\CC)$ on $\Gr_d(\CC^n)$ is transitive which shows that it is transitive on $\Ind_\alpha(K(n))$. As the reflection functors preserve indecomposability and isomorphism classes, by Lemma \ref{compatibleGSigma}, for every $g\in\Gl_n(\CC)$, we obtain a commutative diagram
%\[\xymatrix{\Ind(K(n),\alpha(m,r))\ar[r]^{\Sigma_2}\ar[d]^{g^{-T}}&\Ind(K(n),\alpha(m+1,r))\ar[d]^g\\
%\Ind(K(n),\alpha(m,r))\ar[r]^{\Sigma_2}&\Ind(K(n),\alpha(m+1,r))}\]
%As the $\Gl_n(\CC)$-action is transitive on the left hand side, this implies that it is also transitive on the right hand side.
\end{proof}
\begin{remark}
  There is a more elegant way to prove Proposition \ref{pro:indecomposables} using the notion of stability and moduli spaces.
  Actually, fixing the standard stability induced by the linear form $\Theta(\alpha)=\alpha_2$, it can be shown that all indecomposables are stable and that the moduli space of stable representation is in fact $\Gr_d(\CC^n)$.
  We opted for the proof above because the notion of stability would only be used at this point and we wanted to keep the exposition as simple as possible. 
\end{remark}
